Social Network Analysis: Week 2: Basic Network Structure

### 2022-02-25 docs/3. SNA - Assignments/sna_Week3/DRAFT-week3-Assignment.Rmd

# Week 2: Scripts
1.  [**Identify an existing data set.**]{.ul}
```{r Set Working Directory, include=FALSE}
getwd()

setwd("C:/Users/white/iCloudDrive/Documents/1_DACSS_UMassAmherst/Dacss_RProjects/3. Sp22_rProjects/Dacss_679- Coursework/GIT-ajw_Learning-SocialNetworkAnalysis/docs/3. SNA - Assignments")

@@

```{r Read raw data into R}

library(readr)

index_by_name_Smuggler_Cove_cocktail_recipies <- read_csv("C:/Users/white/iCloudDrive/Documents/1_DACSS_UMassAmherst/Dacss_RProjects/3. Sp22_rProjects/Dacss_679- Coursework/GIT-ajw_Learning-SocialNetworkAnalysis/docs/3. SNA-Data/index-by_name-Smuggler_Cove-cocktail_recipies.csv", 
    col_names = FALSE, col_types = cols(X1 = col_skip(), 
        X4 = col_skip()), na = "empty")

View(index_by_name_Smuggler_Cove_cocktail_recipies)

summary(index_by_name_Smuggler_Cove_cocktail_recipies)

 # Smuggler's Cove

@@

2. Briefly describe the network dataset.
```{r Identify critical features of data:, echo=FALSE}

dim(smugDrink_Indx.df)
colnames(smugDrink_Indx.df)

@@

-   *Be sure to provide information about network size (e.g., information obtained from network description using week 1 network basics tutorial commands.)*

1.  *Identify and describe content of nodes and links,*

-   each of the 16 Florentine Family Names represent a Node in the network.

```{r Colnames}

@@

2.  *Identify format of data set (i.e., matrix or edgelist, directed or not, weighted or not), and*

Vertex attribute names: vertex.names

```{r Dimmentions, echo=TRUE}

  dim(smugDrink_Indx.df)

@@

```{r Network Attributes, echo=TRUE}

  network(smugDrink_Indx.df)

@@

3.  *Identify whether attribute data are present.*

```{r Summary,echo=TRUE}

  summary(smugDrink_Indx.df,
          print.adj = TRUE
          )
```

## Create & Describe: Directed Adjacency Matrix

*Florentine Families Directed Adjacency Matrix Network Objects:*
    **iGraph** smugDrink_Indx.df.ig 
    **Statnet** smugDrink_Indx.df.stat

iGraph and Statnet both use {r graph4} for Directed Adjacency Matrix Objects

```{r graph4, statnet: Directed: Adjacency Network, echo=TRUE}

    smugDrink_Indx.df.stat<-
        network(
        smugDrink_Indx.df,
        directed=F,
        matrix.type="adjacency"
        )

#     provides a description of several critical network features
    print(flo.stat)

@@

```{r graph4, igraph: Directed: Adjacency Network, echo=TRUE}

    flo.ig<-
      graph.adjacency(
      flo,
      mode="undirected"
      ) 
    print(flo.ig)
```

```{r igraph: Network Description, echo=TRUE}

#     Count Vertices
    vcount(flo.ig)

#     Count Edges
    ecount(flo.ig)

# Is this a Bipartite or single mode network?
    is_bipartite(flo.ig)

#  Are edges directed or undirected?
    is_directed(flo.ig)

#Are edges weighted or unweighted?
    is_weighted(flo.ig)

```
 **Vertex and Edge Attributes**

```{r statnet: Attribute Data: Directed: Adjacency Network, echo=TRUE}

#     access vertex attributes
  head(flo.stat %v% "vertex.names")

#      list the names of vertex attributes
  network::list.vertex.attributes(flo.stat)

#     access edge attribute
  head(flo.stat%e% "weight")

#  Generate a list the names of edge attributes
  network::list.edge.attributes(flo.stat)

#   summarize numeric network attribute
  summary(flo.stat  %v% "name")

```{r igraph: Attribute Data: Directed: Adjacency Network, echo=TRUE}

#     access vertex attribute 
  V(flo.ig)$name
#     Generate a list the names of vertex attributes
  igraph::vertex_attr_names(flo.ig)

#     access edge attribute
  E(flo.ig)$weight

#   Generate a list the names of edge attributes  
  igraph::edge_attr_names(flo.ig)

#   summarize numeric network attribute
  summary(E(flo.ig)$weight)

# 3. Explore the dataset using commands from week 2 tutorial.

5.  *geodesic and path distances,*

## Path Length and Geodesic

```{r iGraph:Nodes: Path: Length, echo=TRUE}

#     Calculate distances between two nodes
    distances(flo.ig,"Bischeri","Ridolfi")

#     Calculate distance between two nodes using unweighted edges
    distances(flo.ig,"Bischeri", "Castellani",weights=NA)

We can also find all of the shortest paths between two famalies in this network.
These list the starting node and the ending node with all nodes inbetween for each path. 
The option weights=NA means that any available edge weights are ignored.

@@

#     isolating the distances between specific node
    all_shortest_paths(flo.ig,"Strozzi","Tornabuoni", weights=NA)$res
## Note: manually tell igraph to ignore edge weights.
```

The concept of shortest path to describe the overall network structure can be more useful

```{r iGraph:Network: Path: Shortest, echo=TRUE}

#     find average shortest path for network
    average.path.length(flo.ig,directed=F)

@@

Classifies all dyads in the network as: - Reciprocal (mutual), or mut - Asymmetric (non-mutual), or asym, and - Absent, or null

```{r statnet: Dyad Census , echo=TRUE}

    sna::dyad.census(flo.stat)

```{r iGraph: Dyad Census, echo=TRUE}

  igraph::dyad.census(flo.ig)

@@

Classifies all triads in the network. The triad census provides a fundamental descriptive insight into the types of triads found in a particular dataset.

```{r statnet: Triad Census, echo=TRUE}

#     Classify all Triad in the network: Directed Network
  sna::triad.census(flo.stat, mode= )

#     Classify all Triad in the network: Undirected Network
   sna::triad.census(flo.stat, mode="graph")

#     total number of all four triad types returned by triad census
   sum(sna::triad.census(flo.stat, mode="graph"))

@@

```{r iGraph: Triad Census, echo=TRUE}
#Classify all Triad in the network
     igraph::triad_census(flo.ig)

@@

#     network transitivity:
    transitivity(flo.ig)

#     weighted network transitivity:
#   igraph::transivity(flo.ig)

@@

#     Directed Networks transitivity:
  gtrans(flo.stat)

@@

#     Retrive a list of the vertices we are interested
    V(flo.ig)[c("Bischeri" ,
                "Castellani", 
                "Medici",
                "Ridolfi")]

#     check ego network transitivity
transitivity(flo.ig,
             type="local", 
             vids=V(flo.ig)
             [c("Bischeri",
                "Castellani", 
                "Medici",
                "Ridolfi")]) 
#     get global clustering cofficient
  transitivity(flo.ig, type="global")

#     get average local clustering coefficient
  transitivity(flo.ig, type="average")

>>>>

# Week 2: Scripts
## Smuggler's Index: Network Attributes  
  - Edgelist
  - unweghted
  - undirected

Network attributes:

*statnet*
  vertices = 274
  directed = TRUE
  hyper = FALSE
  loops = FALSE
  multiple = FALSE
  bipartite = FALSE
 total edges = 671 
   missing edges = 0 
   non-missing edges = 0 
 density = 0

Edge attributes:
[1]ingTyp:
   character valued attribute

Attribute summary:
  the 10 most common values are:
  JUICE     RUM   SYRUP LIQUEUR BITTERS    SODA   SPICE     GIN  BRANDY   FRUIT 
    135     125     115      84      54      28      19      12       9       7

###   IGRAPH:   Network Description
*igraph*
- Network Size
  - Number Vertex: 274
    - attr: name (v/c)
  - Number Edges: 672
    - attr: ingTyp (e/c)
- Bipartite: FALSE
- Directed: FALSE
- Weighted: FALSE

Names
- Vertex Names: view
- Edge Attribute Names: [1] "ingTyp"
```{r igraph: Network Description, echo=TRUE}

summary(igEL)
#     Count Vertices
    vcount(igEL)

#     Count Edges
    ecount(igEL)

# Is this a Bipartite or single mode network?
    is_bipartite(igEL)

#  Are edges directed or undirected?
    is_directed(igEL)

#Are edges weighted or unweighted?
    is_weighted(igEL)

@@

```{r igraph: Vertex & Attribute info, echo=TRUE}
library(igraph)

#     access vertex attribute 
  head(V(igEL)$name,20)
#     Generate a list the names of vertex attributes
  print(igraph::vertex_attr_names(igEL))

#     access edge attribute
  head(E(igEL)$ingTyp, 20)

#   Generate a list the names of edge attributes  
  igraph::edge_attr_names(igEL)

#   summarize numeric network attribute
  summary(E(igEL)$weight)

@@

###   STATNET:   Network Description 
 *statnet*
  vertices = 274
  directed = TRUE
  hyper = FALSE
  loops = FALSE
  multiple = FALSE
  bipartite = FALSE
 total edges = 671 
   missing edges = 0 
   non-missing edges = 0 
 density = 0

Edge attributes:

 ingTyp:
   character valued attribute
   attribute summary:
   the 10 most common values are:
  JUICE     RUM   SYRUP LIQUEUR BITTERS    SODA   SPICE     GIN  BRANDY   FRUIT 
    135     125     115      84      54      28      19      12       9       7 
```{r statnet: Network Summary & Attribute info, echo=TRUE}

summary(snEL)
print(snEL)
#     access vertex attributes
 head(snEL %v% "vertex.names")

#      list the names of vertex attributes
  network::list.vertex.attributes(snEL)

#     access edge attribute
  head(snEL%e% "ingTyp")

#  Generate a list the names of edge attributes
  network::list.edge.attributes(snEL)

#   summarize numeric network attribute
  summary(snEL  %v% "name")

@@

##  Path Length and Geodesic Distance

- Nodes: 
  - (v/c) Drink Names
  - (v/c) Drink Ingredients
- Links: 
  - (e/c)  Ingredient Type

###  Statnet:   Path Length and Geodesic Distance

    Attribute summary:
  the 10 most common values are:
  JUICE     RUM   SYRUP LIQUEUR BITTERS    SODA   SPICE     GIN  BRANDY   FRUIT 
    135     125     115      84      54      28      19      12       9       7 
```{r STATNET:Path: Length}
library(statnet)

@@

 >head(snEL%e% "ingTyp")

[1] "FRUIT"   "LEAVES"  "JUICE"   "SYRUP"   "LIQUEUR" "RUM"

   >head(snEL %v% "vertex.names")

[1] "(1-INCH-SQUARE) CHUNKS PINEAPPLE" "(1-INCH-SQUARE) PINEAPPLE CHUNKS" "2:1 SIMPLE SYRUP"                
[4] "A WISH FOR GRACE"                 "ABRICOT VIEUX"                    "ABSINTHE BLANC"

###  IGRAPH:   Path Length and Geodesic Distance

```{r iGraph:Nodes: Path: Length, echo=TRUE}
library(igraph)

# Head
head(V(igEL)$name,10)
head(E(igEL)$ingTyp, 10)

#Tail
tail(V(igEL)$name,10)
tail(E(igEL)$ingTyp, 10)

@@

  # 20 Vertex Names

      Head, 10 + Tail, 10
 [1] "AKU AKU"                    
 [2] "(1-INCH-SQUARE) CHUNKS PINEAPPLE"    
 [3] "MINT LEAVES"                
 [4] "LIME JUICE"    
 [5] "DEMERARA SYRUP"             
 [6] "PEACH LIQUEUR (SUCH AS MATHILDE, COMBIER, OR GIFFARD)"
 [7] "BLENDED LIGHTLY AGED RUM"   
 [8] "PUPULE" 
 [9] "ORANGE JUICE"               
 [10] "CINNAMON SYRUP"
 [1] "SINGAPORE SLING"                     
      "BENEDICTINE"
      "HONDO HATTIE'S JUNGLE PUNCH!"    
  [4] "KAHIKO PUNCH"
      "PASSION FRUIT HONEY"                 
      "TRIUMVIRATE PUNCH"   
  [7] "EUREKA PUNCH"                         
      "YULETIDAL WAVE"
      "RUMBUSTION PUNCH"                
 [10] "PUKA PUNCH"

  The shortest path to describe the overall network structure 
```{r iGraph:Network: Path: Shortest, echo=TRUE}

#     find average shortest path for network
    average.path.length(igEL,directed=F)

@@

```{r iGraph:Nodes: Path: Length, echo=TRUE}
library(igraph)

#     Calculate distance between two nodes using unweighted edges

    distances(igEL,"AKU AKU" ,  "SINGAPORE SLING",weights=NA)
    distances(igEL,"AKU AKU" ,  "HONDO HATTIE'S JUNGLE PUNCH!",weights=NA)
    distances(igEL,"AKU AKU" ,"KAHIKO PUNCH" ,weights=NA)
    distances(igEL,"AKU AKU" , "PASSION FRUIT HONEY"   ,weights=NA)
    distances(igEL,"AKU AKU" , "TRIUMVIRATE PUNCH"   ,weights=NA)
    distances(igEL,"AKU AKU" ,"EUREKA PUNCH" ,weights=NA)
    distances(igEL,"AKU AKU" , "YULETIDAL WAVE" ,weights=NA)
    distances(igEL,"AKU AKU" ,"RUMBUSTION PUNCH"   ,weights=NA)
    distances(igEL,"AKU AKU" ,"PUKA PUNCH",weights=NA)

@@

Find *all* of the shortest paths between two specific nodes in this network.

@@

#     isolating the distances between specific node

## Note: manually tell igraph to ignore edge weights.

all_shortest_paths(igEL,
                     "AKU AKU" ,
                     "SINGAPORE SLING",
                     weights=NA)$res

all_shortest_paths(igEL,
                   "AKU AKU",
                   "HONDO HATTIE'S JUNGLE PUNCH!",
                   weights=NA)$res

all_shortest_paths(igEL,
                   "AKU AKU",
                   "KAHIKO PUNCH",
                   weights=NA)$res

all_shortest_paths(igEL,
                     "AKU AKU", 
                     "PASSION FRUIT HONEY",
                     weights=NA)$res

all_shortest_paths(igEL,
                   "AKU AKU",
                   "TRIUMVIRATE PUNCH",
                   weights=NA)$res

  all_shortest_paths(igEL,"AKU AKU",
                     "EUREKA PUNCH",
                     weights=NA)$res

  all_shortest_paths(igEL,"AKU AKU", 
                     "YULETIDAL WAVE" ,
                     weights=NA)$res

  all_shortest_paths(igEL,
                     "AKU AKU" ,
                     "RUMBUSTION PUNCH",
                     weights=NA)$res

  all_shortest_paths(igEL,
                     "AKU AKU" ,
                     "PUKA PUNCH",
                     weights=NA)$res
```

@@

Classifies all dyads in the network as:
- Mutual - Reciprocal (mutual) 
- Asymmetric -  (non-mutual), 
- Null - Absent

```{r statnet: Dyad Census , echo=TRUE}
library(statnet)
    sna::dyad.census(snEL)

```
```{r iGraph: Dyad Census, echo=TRUE}
library(igraph)

dyad.census(igEL)

@@

Classifies all triads in the network.

```{r statnet: Triad Census, echo=TRUE}
library(sna)
#     Classify all Triad in the network:

#Directed Network
#**triad.census(snEL, mode="digraph")*

#     Classify all Triad in the network: Undirected Network
   sna::triad.census(
     snEL,
     mode="graph")

#     total number of all four triad types returned by triad census
   sum(
     sna::triad.census(
       snEL,
       mode="graph")
     )

#Davis and Leinhardt classification of the triad classification scheme 
   # describes each triad by a string of four elements: 
   # the number of mutual (complete) dyads within the triad; 
   # the number of asymmetric dyads within the triad; 
   # the number of null (empty) dyads within the triad; 
   #and a configuration code for the triads which are not uniquely distinguished by the first three distinctions.

   sna::triad.classify(
     snEL,
     tri=c(1,2,3),
     mode==c("digraph")
   )
  #Classify the triads (1,2,3) and (2,3,4)

   triad.classify(snEL,tri=c(1:14))
   gplot(snEL,c[1:14])

@@

```{r iGraph: Triad Census, echo=TRUE}
library(igraph)
#Classify all Triad in the network
     igraph::triad_census(igEL)

@@

#     network transitivity:
    transitivity(igEL)

#     weighted network transitivity:
#   igraph::transivity(igEL)

@@

#     Directed Networks transitivity:
  gtrans(snEL)

@@

#     Retrive a list of the vertices we are interested
    V(igEL)[c("" ,
                "", 
                "Medici",
                "")]

#     check ego network transitivity
transitivity(igEL,
             type="local", 
             vids=V(igEL)
             [c("",
                "", 
                "Medici",
                "")]) 
#     get global clustering cofficient
  transitivity(igEL, type="global")

#     get average local clustering coefficient
  transitivity(igEL, type="average")

### 2022-02-25 docs/3. SNA - Assignments/week2-Assignment-Basic_Network_Structure.Rmd

#     Retrive a list of the vertices we are interested
    V(flo.ig)[c("Bischeri" ,
                "Castellani", 
                "Medici",
                "Ridolfi")]

#     check ego network transitivity
transitivity(flo.ig,
             type="local", 
             vids=V(flo.ig)
             [c("Bischeri",
                "Castellani", 
                "Medici",
                "Ridolfi")]) 
#     get global clustering cofficient
  transitivity(flo.ig, type="global")

>>>>

#     Retrive a list of the vertices we are interested
    V(flo.ig)[c("Bischeri" ,
                "Castellani", 
                "Medici",
                "Ridolfi")]
#     check ego network transitivity
transitivity(flo.ig,
             type="local", 
             vids=V(flo.ig)
             [c("Bischeri",
                "Castellani", 
                "Medici",
                "Ridolfi")]) 
#     get global clustering cofficient
  transitivity(flo.ig, type="global")
